### 2024-10-31 manuscript.tex
Update on Overleaf.

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is a 5-km Icosahedral Nonhydrostatic Weather and Climate Model
(ICON). Due to the resolution, parametrisations of convection and clouds are
avoided. Standard-resolution models have substantial cloud biases over the
Southern Ocean (SO), affecting radiation and sea surface temperature.  We
evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses.  The SO is
dominated by low clouds, which cannot be observed properly from space due to
overlapping clouds, attenuation, and ground clutter.  Instead, we analysed
about 2400 days of lidar observations from 31 voyages and a station using a
ground-based lidar simulator.  ICON and the reanalyses underestimate the total
cloud fraction by about 10 and 20\%, respectively. ICON and ERA5 overestimate
the cloud occurrence peak at about 500 m, potentially explained by their
lifting condensation level being too high.  The reanalyses strongly
underestimate near-surface clouds or fog.  MERRA-2 tends to underestimate cloud
occurrence at all heights. Unstable conditions are the most problematic for the
models. In daily cloud cover, ICON and the reanalyses tend to be about 1 and 2
oktas clearer, respectively. Compared to radiosondes, potential temperature is
accurate in all, but ICON is too unstable over low-latitude SO and too humid in
the boundary layer. MERRA-2 is too humid at all heights. SO cloud biases are a
substantial issue in the GSRM, but it is an improvement over the
lower-resolution reanalyses. Explicitly resolved convection and cloud processes
were not enough to address the model cloud biases.

@@

Increasing climate model resolution is a common way of improving model accuracy
\citep{mauritsen2022}. This is of course limited by the available computational
power and a tradeoff with increasing parametrisation complexity. Current
computational availability and acceleration from general-purpose computing on
graphics processing units (GPGPUs) is progressing to enable km-scale (also
called k-scale) Earth system models (ESMs) and coupled atmosphere--ocean
general circulation models (AOGCMs) in research conditions today and
operationally in the forthcoming years. Therefore, it represents a natural
advancement in climate modeling.  Global strom-resolving models (GSRMs) are
emerging as a new front in the development of high-resolution global climate
models, with horizontal grid resolutions of about 2--8 km
\citep{satoh2019,stevens2019}. This is enough to resolve mesoscale convective
storms, but smaller-scale convective plumes and smaller-scale cloud structure
remain unresolved. At a 5-km scale, non-hydrostatic processes also become
important, and for this reason such models are generally non-hydrostatic. The
terms global cloud-resolving models or global convection-permitting/-resolving
models are also sometimes used interchangeably with GSRMs but imply that clouds
or convection are resolved explicitly, which is not entirely true for GSRMs, as
this would require an even higher horizontal resolution \citep{satoh2019}.
Representative of these efforts is the DYnamics of the Atmospheric general
circulation Modeled On Non-hydrostatic Domains (DYAMOND) project
\citep{stevens2019,dyamond}, which is an intercomparison of nine global GSRMs
on two 40-day time periods in summer (1 August -- 10 September 2016) and winter
(20 January -- 1 March 2020). A new one-year GSRM intercomparison is currently
proposed by \cite{takasuka2024}, with the hope of also evaluating the seasonal
cycle and large-scale circulation.  An alternative to using a computationally
costly GSRM is to train an artificial neural network on GSRM output and use it
for subgrid-scale clouds, as done with the GSRM ICON by \cite{grundner2022} and
\cite{grundner2023}.

@@

Multiple cloud properties have an effect on the shortwave (SW) and longwave
(LW) radiation. On the first order, it is the total cloud fraction, cloud
phase, and the liquid and ice water path. These are influenced by the
atmopheric thermodynamics, convection and circulation, and indirect and direct
effects of aerosols. On smaller orders, it is the cloud droplet size
distribution, ice crystal habit, cloud lifetime, and direct radiative
interaction with aerosols.  In the 6$^\mathrm{th}$ phase of the Coupled Model
Intercomparison Project [CMIP6; \cite{eyring2016}], the cloud feedback has
increased relative to CMIP5 \citep{zelinka2020}, also being one of the main
reasons for the higher climate sensitivity of CMIP6 models.

The Southern Ocean (SO) is known to be a problematic region for climate model
biases due to reasons such as a lack of surface and in situ observations and
being a lower priority region for numerical weather prediction (NWP) and
climate model development because it is far away from populated areas of
interest.  Nevertheless, radiation biases and changes over an area of its size
have a very substantial influence on the global climate, and the SO is an
important part of the global ocean conveyor belt.  Marine clouds have
a disproportionate effect on top of atmosphere (TOA) SW radiation due to the
relatively low albedo of the sea surface.  The relative longitudinal symmetry
of the SO means that model cloud biases tend to be similar across longitudes.
Here, we conventionally refer to the SO as ocean regions south of 40°S,
low-latitude SO as 40--55°S and high-latitude SO as south of 55°S.

SO radiation biases have been relatively large and systematic compared to the
rest of the globe since at least CMIP3 \citep{trenberth2010}, and the SO SW
cloud radiative effect (CRE) bias is still positive in eight analysed CMIP6
models analysed by \cite{schuddeboom2021} over high-latitude SO, whereas over
low-latitude SO it tends to be more neutral or negative in some models. Too
much absorbed SW radiation over the SO was also identified in the GSRM SCREAM
\cite{caldwell2021}. Compensating biases are possible, such as the `too few too
bright' cloud bias, characterised by too small cloud fraction and too large
cloud albedo \citep{wall2017,kuma2020}, previously described by
\cite{webb2001}, \cite{weare2004}, \cite{zhang2005}, \cite{karlsson2008},
\cite{nam2012}, \cite{klein2013}, and \cite{bender2017} in other regions and
models. That is, a model maintains a resonable SW radiation balance by
reflecting too much SW radiation by cloudy sky but has too little cloudy sky
area overall. A study by \cite{konsta2022} showed that this type bias is still
present in six analysed CMIP6 models in tropical marine clouds, using the
GCM-Oriented Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation
(CALIPSO) Cloud Product [CALIPSO--GOCCP; \cite{chepfer2010}] and Polarization
\& Anisotropy of Reflectances for Atmospheric Sciences coupled with
Observations from a Lidar [PARASOL; \cite{lier2008}] as a reference. They
suggest improper simulation of subgrid-scale cloud heterogeneity as a cause.
Compensating cloud biases in the Australian Community Climate and Earth System
Simulator (ACCESS) – Atmosphere-only model version 2 (AM2) over the SO were
analysed by \cite{fiddes2022} and \cite{fiddes2024}.  \cite{possner2022} showed
that over the SO, the DYAMOND GSRM ICON underestimates low-level cloud fraction
on the order of 30\% [relative to the Moderate Resolution Imaging
Spectroradiometer (MODIS)] and overestimates downwelling TOA SW radiation on
the order of 10 Wm$^\mathrm{-2}$ in  [relative to the Clouds and the Earth’s
Radiant Energy System (CERES)] in the highest model resolution run (2.5 km).
\cite{zhao2022} reported a similar SW radiation bias in five analysed CMIP6
models over high-latitude SO and the total cloud fraction underestimation on
the order of 10\% over the entire 40--60°S SO. Recently, \cite{ramadoss2024}
analysed 48 hours of km-scale ICON limited area model NWP simulation over a SO
region adjacent to Tasmania agaist the Clouds, Aerosols, Precipitation,
Radiation, and atmospherIc Composition Over the southeRn oceaN (CAPRICORN)
voyage cloud and precipitation observations. They found the ICON cloud optical
thickness underestimated relative to Himawari‐8 satellite observations, but
also large differences in cloud top phase.

In general, sea surface temperature (SST) biases in the SO can originate either
in the atmosphere, caused by too much shortwave heating of the surface, too
little longwave cooling of the surface, or in the ocean circulation.
Interactions of both are also possible, due to, for example SST affecting
clouds and clouds affecting the surface radiation.  \cite{zhang2023} have shown
that SST biases have improved in CMIP6 compared to CMIP5 [relative to the
European Centre for Medium-Range Weather Forecasts (ECMWF) Reanalysis 5; ERA5],
with SST overall increasing in the later CMIP phase. However, over the SO this
resulted in an even higher positive bias, especially in the Atlantic Ocean (AO)
sector of the SO, increasing by up to 1°C.  \cite{luo2023} identified that the
SO SST bias in an ensemble of 18 CMIP6 models originates not from the surface
heat and radiation fluxes (using reanalyses as a reference), but from a warm
bias in the Northern Atlantic Deep Water.

>>>>

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is the 5-km Icosahedral Nonhydrostatic Weather and Climate Model
(ICON). Due to the resolution, parametrisations of convection and clouds are
avoided. Standard-resolution models have substantial cloud biases over the
Southern Ocean (SO), affecting radiation and sea surface temperature.  We
evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses.  The SO is
dominated by low clouds, which cannot be observed accurately from space due to
overlapping clouds, attenuation, and ground clutter.  Instead, we analysed
about 2400 days of lidar observations from 31 voyages and a station using a
ground-based lidar simulator.  ICON and the reanalyses underestimate the total
cloud fraction by about 10 and 20\%, respectively. ICON and ERA5 overestimate
the cloud occurrence peak at about 500 m, potentially explained by their
lifting condensation levels being too high.  The reanalyses strongly
underestimate near-surface clouds or fog.  MERRA-2 tends to underestimate cloud
occurrence at all heights. Unstable conditions are the most problematic for the
models. In daily cloud cover, ICON and the reanalyses tend to be about 1 and 2
oktas clearer, respectively. Compared to radiosondes, potential temperature is
accurate in all, but ICON is too unstable over the low-latitude SO and too humid in
the boundary layer. MERRA-2 is too humid at all heights. SO cloud biases are a
substantial issue in the GSRM, but are an improvement over the
lower-resolution reanalyses. Explicitly resolved convection and cloud processes
were not enough to address the model cloud biases.

@@

Increasing climate model resolution is a common way of improving model accuracy
\citep{mauritsen2022}. This is of course limited by the available computational
power and a tradeoff with increasing parametrisation complexity. Current
computational availability and acceleration from general-purpose computing on
graphics processing units (GPUs) is progressing to enable km-scale (also
called k-scale) Earth system models (ESMs) and coupled atmosphere--ocean
general circulation models (AOGCMs) in research conditions today and
operationally in the forthcoming years. Therefore, it represents a natural
advance in climate modelling.  Global storm-resolving models (GSRMs) are
emerging as a new front in the development of high-resolution global climate
models, with horizontal grid resolutions of about 2--8 km
\citep{satoh2019,stevens2019}. This is enough to resolve mesoscale convective
storms, but smaller-scale convective plumes and smaller-scale cloud structure
remain unresolved. At a 5-km scale, non-hydrostatic processes also become
important, and for this reason such models are generally non-hydrostatic. The
terms global cloud-resolving models or global convection-permitting/-resolving
models are also sometimes used interchangeably with GSRMs but imply that clouds
or convection are resolved explicitly, which is not entirely true for GSRMs, as
this would require an even higher horizontal resolution \citep{satoh2019}.
Representative of these efforts is the DYnamics of the Atmospheric general
circulation Modeled On Non-hydrostatic Domains (DYAMOND) project
\citep{stevens2019,dyamond}, which is an intercomparison of nine global GSRMs
over two 40-day time periods in summer (1 August -- 10 September 2016) and winter
(20 January -- 1 March 2020). A new one-year GSRM intercomparison is currently
proposed by \cite{takasuka2024}, with the hope of also evaluating the seasonal
cycle and large-scale circulation.  An alternative to using a computationally
costly GSRM is to train an artificial neural network on GSRM output and use it
for subgrid-scale clouds, as done with the GSRM ICON by \cite{grundner2022} and
\cite{grundner2023}.

@@

Multiple cloud properties have an effect on the shortwave (SW) and longwave
(LW) radiation. To first order, the total cloud fraction, cloud
phase, and the liquid and ice water path are the most important cloud properties influencing SW and LW radiation. These properties are  in turn influenced by the
atmospheric thermodynamics, convection and circulation, and indirect and direct
effects of aerosols. Second order effects on SW and LW radiation are associated with the cloud droplet size
distribution, ice crystal habit, cloud lifetime, and direct radiative
interaction with aerosols.  In the 6$^\mathrm{th}$ phase of the Coupled Model
Intercomparison Project [CMIP6; \cite{eyring2016}], the cloud feedback has
increased relative to CMIP5 \citep{zelinka2020}, this being one of the main
reasons for the higher climate sensitivity of CMIP6 models.

The Southern Ocean (SO) is known to be a problematic region for climate model
biases due to a lack of surface and in situ observations and
being a lower priority region for numerical weather prediction (NWP) and
climate model development because of its distance from populated areas of
interest.  Nevertheless, radiation biases and changes over an area of its size
have a very substantial influence on the global climate, and the SO is an
important part of the global ocean conveyor belt.  Marine clouds have
a disproportionate effect on top of atmosphere (TOA) SW radiation due to the
relatively low albedo of the sea surface.  The relative longitudinal symmetry
of the SO means that model cloud biases tend to be similar across longitudes.
Here, we conventionally refer to the SO as ocean regions south of 40°S,
low-latitude SO as 40--55°S and high-latitude SO as south of 55°S.

SO radiation biases have been relatively large and systematic compared to the
rest of the globe since at least CMIP3 \citep{trenberth2010}, and the SO SW
cloud radiative effect (CRE) bias is still positive in eight analysed CMIP6
models analysed by \cite{schuddeboom2021} over the high-latitude SO, whereas over
the low-latitude SO it tends to be more neutral or negative in some models. Too
much absorbed SW radiation over the SO was also identified in the GSRM SCREAM
\cite{caldwell2021}. Compensating biases are possible, such as the `too few too
bright' cloud bias, characterised by too small cloud fraction and too large
cloud albedo \citep{wall2017,kuma2020}, previously described by
\cite{webb2001}, \cite{weare2004}, \cite{zhang2005}, \cite{karlsson2008},
\cite{nam2012}, \cite{klein2013}, and \cite{bender2017} in other regions and
models. That is, a model maintains a reasonable SW radiation balance by
reflecting too much SW radiation from clouds, but has too little cloud
area overall. A study by \cite{konsta2022} showed that this type of bias is still
present in six analysed CMIP6 models in tropical marine clouds, using the
GCM-Oriented Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation
(CALIPSO) Cloud Product [CALIPSO--GOCCP; \cite{chepfer2010}] and Polarization
\& Anisotropy of Reflectances for Atmospheric Sciences coupled with
Observations from a Lidar [PARASOL; \cite{lier2008}] as a reference. They
suggest improper simulation of subgrid-scale cloud heterogeneity as a cause.
Compensating cloud biases in the Australian Community Climate and Earth System
Simulator (ACCESS) – Atmosphere-only model version 2 (AM2) over the SO were
analysed by \cite{fiddes2022} and \cite{fiddes2024}.  \cite{possner2022} showed
that over the SO, the DYAMOND GSRM ICON underestimates low-level cloud fraction
on the order of 30\% [relative to Moderate Resolution Imaging
Spectroradiometer (MODIS) data] and overestimates downwelling TOA SW radiation on
by approximately 10 Wm$^\mathrm{-2}$ in  [relative to the Clouds and the Earth’s
Radiant Energy System (CERES) data] in the highest model resolution run (2.5 km).
\cite{zhao2022} reported a similar SW radiation bias in five analysed CMIP6
models over high-latitude SO and the total cloud fraction underestimation on
the order of 10\% over the entire 40--60°S SO. Recently, \cite{ramadoss2024}
analysed 48 hours of km-scale ICON limited area model NWP simulations over a SO
region adjacent to Tasmania against the Clouds, Aerosols, Precipitation,
Radiation, and atmospherIc Composition Over the southeRn oceaN (CAPRICORN)
voyage cloud and precipitation observations \cite{mcfarquhar2021}. They found the ICON cloud optical
thickness was underestimated relative to Himawari‐8 satellite observations, but
also identified large differences in cloud top phase.

In general, sea surface temperature (SST) biases in the SO can originate either
in the atmosphere, caused by too much shortwave heating of the surface, too
little longwave cooling of the surface, or in the ocean circulation.
Interactions of both are also possible, due to, for example SST affecting
clouds and clouds affecting the surface radiation. \cite{zhang2023} has shown
that SST biases have improved in CMIP6 compared to CMIP5 [relative to the
European Centre for Medium-Range Weather Forecasts (ECMWF) Reanalysis 5; ERA5],
with SST overall increasing in the later CMIP phase. However, over the SO this
resulted in an even higher positive bias, especially in the Atlantic Ocean (AO)
sector of the SO, increasing by up to 1°C.  \cite{luo2023} identified that the
SO SST bias in an ensemble of 18 CMIP6 models originates not from the surface
heat and radiation fluxes (using reanalyses as a reference), but from a warm
bias in the Northern Atlantic Deep Water.
